Language-Based Security for Low-Level MPC

### 2024-05-21 ppdp24/example-gmw.tex
checkpoint

To properly integrate the local reasoning of Lemma \ref{lemma-gmwtactic} with
the global reasoning of Lemma \ref{lemma-gmwinvariant}, we can demonstrate
the following.
\begin{lemma}
  \label{lemma-conditioning}
  Give $\pmf$, each of the following hold for $S,V_1,V_2,V_3 \in \dom(\pmf)$:
  \begin{enumerate}
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_2}{\detx{V_2}}$, then $\condp{\pmf}{S}{\detx{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_1}{\unix{V_2}}$, then $\condp{\pmf}{S}{\unix{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condsep{\pmf}{V_1}{V_2}{V_3}$, then $\condsep{\pmf}{S}{V_2}{V_3}$.
    %\item Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$
    %  and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
    %\item Given $\condsep{\pmf}{X_1 \cup X_2}{X_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
    %  then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
  \end{enumerate}
\end{lemma}

Then we can put the pieces together to prove the invariant, using automated tactics
for gate-level reasoning.  
\begin{proof}[Proof of Lemma \ref{lemma-gmwinvariant}]
  By induction on the length of $(\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}]))$.
  Encoding establishes the invariant in the basis. The most interesting inductive
  case is the $\andgmw$ gate. 
  \paragraph{Case $\andgmw$.} In this case we have:
  $$
  \begin{array}{c}
  (\varnothing,e) \redxs (\prog',\decodegmw(E[\andgmw(z,\mesg{x},\mesg{y})])) \redxs \\
    (\prog,\decodegmw(E[\mesg{z}]))
  \end{array}
  $$

  \begin{enumerate}[\hspace{5mm}(x.1)]
  \item $\conddetx{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}$
  \item $\condunix{\progtt(\prog)}{S^1}{M^1_C}$
  %\item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
  \end{enumerate}

  \begin{enumerate}[\hspace{5mm}(y.1)]
  \item $\conddetx{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}$
  \item $\condunix{\progtt(\prog)}{S^2}{M^2_C}$
  %\item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
  \end{enumerate}

  \begin{enumerate}[\hspace{5mm}(z.1)]
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}

  Given $x.1$ and $y.1$ and assumed uniqueness of random variables
  used in each gate we have $\conddetx{\progtt(\prog)}{S}{\{
    \macgv{\mesg{x}},\macgv{\mesg{y}}\}}$, so it follows from $z.1$
  and Lemma \ref{lemma-conditioning} (1) that
  $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$, and also by
  $z.3$ and Lemma \ref{lemma-conditioning} (3):
  $$\condsep
  {\progtt(\prog)}
  {S}
  {\{ \macgv{\mesg{z}} \}}
  {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$$
\end{proof}

@@

\begin{proof}
  Given that $e$ is a GMW circuit protocol, then by definition we have:
  $$
  (\varnothing,e) \redxs (\prog',\decodegmw(\mesg{z})) \redxs (\prog,())
  $$
  where by Lemma \ref{lemma-gmwinvariant} and definition of $\decodegmw$,
  for any $H$ and $C$ letting $\iov(\prog) = S \cup M \cup P \cup O$ we
  have:
  \begin{mathpar}
    \conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}

    \condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{(M_C \cup \{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
  \end{mathpar}
  and by definition of $\decodegmw$ the value $\{\macgv{\mesg{z}}\}$ is output
  after the parties publicly reveal $\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \}$, so we have:
  \begin{mathpar}
    \conddetx{\progtt(\prog)}{S_H \cup S_C}{O}

    \condsep{\progtt(\prog)}{S_H \cup S_C}{O}{(M_C \cup P)}
  \end{mathpar}
  and thus by Lemma \ref{lemma-conditioning} (4) we have:
  $$
  \condsep{\progtt(\prog)}{S_C}{(O \cup S_H)}{(M_C \cup P)}
  $$
  which implies the result.
\end{proof}

>>>>

To properly integrate the local reasoning of Lemma \ref{lemma-gmwtactic} with
the global reasoning of Lemma \ref{lemma-gmwinvariant}, we can demonstrate
the following.
\begin{lemma}
  \label{lemma-conditioning}
  Give $\pmf$, each of the following hold for $S,V_1,V_2,V_3 \in \dom(\pmf)$:
  \begin{enumerate}
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_2}{\detx{V_2}}$, then $\condp{\pmf}{S}{\detx{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_1}{\unix{V_2}}$, then $\condp{\pmf}{S}{\unix{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condsep{\pmf}{V_1}{V_2}{V_3}$, then $\condsep{\pmf}{S}{V_2}{V_3}$.
    %\item Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$
    %  and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
    %\item Given $\condsep{\pmf}{S_1 \cup S_2}{V_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
    %  then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
  \end{enumerate}
\end{lemma}

Then we can put the pieces together to prove the invariant, using automated tactics
for gate-level reasoning.  
\begin{proof}[Proof of Lemma \ref{lemma-gmwinvariant}]
  By induction on the length of $(\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}]))$.
  Encoding establishes the invariant in the basis. The most interesting inductive
  case is the $\andgmw$ gate. 
  \paragraph{Case $\andgmw$.} In this case we have:
  $$
  \begin{array}{c}
  (\varnothing,e) \redxs (\prog',\decodegmw(E[\andgmw(z,\mesg{x},\mesg{y})])) \redxs \\
    (\prog,\decodegmw(E[\mesg{z}]))
  \end{array}
  $$
  Letting $\iov(\prog) = S \cup M$, by definition and the induction hypothesis there
  exist $S^1,S^2 \subseteq S$ and $M^1,M^2 \subseteq M$
  such that:
  \begin{enumerate}[\hspace{5mm}(x.1)]
  \item $\conddetx{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}$
  \item $\condunix{\progtt(\prog)}{S^1}{M^1_C}$
  \item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
  \end{enumerate}
  and also:
  \begin{enumerate}[\hspace{5mm}(y.1)]
  \item $\conddetx{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}$
  \item $\condunix{\progtt(\prog)}{S^2}{M^2_C}$
  \item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
  \end{enumerate}
  and by Lemma \ref{lemma-gmwtactic} we have:
  \begin{enumerate}[\hspace{5mm}(z.1)]
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}
  Given $x.1$ and $y.1$ and assumed uniqueness of random variables
  used in each gate we have $\conddetx{\progtt(\prog)}{S}{\{
    \macgv{\mesg{x}},\macgv{\mesg{y}}\}}$, so it follows from $z.1$
  and Lemma \ref{lemma-conditioning} (1) that
  $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$, and also by
  $z.3$ and Lemma \ref{lemma-conditioning} (3):
  $$\condsep
  {\progtt(\prog)}
  {S}
  {\{ \macgv{\mesg{z}} \}}
  {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$$
  Additionally Lemma \ref{lemma-conditioning} (2) gives $\condunix{\progtt(\prog)}{S}{\{ \elab{\mesg{z}}{\cid} \}}$,
  and $x.2$ and $y.2$ give $\condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C}$. Facts $x.3$ and $y.3$ together with
  uniqueness of gate identifiers implies:
  $$
  \condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C \cup \{  \elab{\mesg{z}}{\cid} \}}
  $$
  This implies the result.
\end{proof}

@@

\begin{proof}
  Given that $e$ is a GMW circuit protocol, then by definition we have:
  $$
  (\varnothing,e) \redxs (\prog',\decodegmw(\mesg{z})) \redxs (\prog,())
  $$
  where by Lemma \ref{lemma-gmwinvariant} and definition of $\decodegmw$,
  for any $H$ and $C$ letting $\iov(\prog) = S \cup M \cup P \cup O$ we
  have:
  \begin{mathpar}
    \conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}
    \condunix{\progtt(\prog)}{S}{M_C}
    \condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}
  \end{mathpar}
  and by definition of $\decodegmw$ the value $\{\macgv{\mesg{z}}\}$ is output
  after the parties publicly reveal $\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \}$, so we have:
  \begin{mathpar} 
    \condunix{\progtt(\prog)}{S}{M_C}

    \condsep{\progtt(\prog)}{S}{O}{P}
  \end{mathpar}
  which implies the result.
\end{proof}
